A universal of human social cognition: Children from 17 communities process gaze in similar ways

### 2024-04-03 supplement/supplement.Rmd
update paper for submission

---
title: "Unity in diversity: Children from 17 diverse communities process gaze in similar ways"
subtitle: "Supplemental material"
author: "..."
output: 
  bookdown::pdf_document2:
    toc: yes 
    number_sections: false    
  bookdown::html_document2:
    toc: yes
    toc_float: true
    fig_caption: yes
    #code_folding: hide
    number_sections: false
bibliography: ../paper/library.bib
csl: science.csl
header-includes:
  \usepackage{caption}
  \renewcommand{\figurename}{Supplementary Figure}
  \renewcommand{\tablename}{Supplementary Table}
---

@@

# Overview

This docuemnt provides a detailed overview of the participants and methods of the in the section [site descriptions](#site-descriptions)

@@

A total of `r sum(as.numeric(demtab$N))` children between `r round(min(data$ageinyears), 2)` and `r round(max(data$ageinyears), 2)` provided data for the study. Children lived in `r n_distinct(data$community)` different communities, located in `r n_distinct(data$country)` different countries. Table \@ref(tab:demtable) gives the sample size per community together with some basic demographic information. Figure \@ref(fig:figpred1)A shows the age distribution in each setting. For some children, the exact birthday was unknown. In such cases, we set the birthday to the 30th of June of the year that would make them fall into the reported age category. Additional children were recruited but did not contribute data to the study. The recruitment strategy varied from community to community and we therefore cannot provide detailed information about the number of children that were excluded. All children who contributed at least four valid data points were included in the final sample, however, we excluded children when parents did not fill out the questionnaire correctly (e.g., when the number of children living in the household was reported to be larger than the household itself or when the number of children reported to live in the household equaled the number of children younger than the child being tested). Twelve children had to be excluded for one of these reasons (one from Akure, two from Stanford, seven from Pune and two from Windhoek). We did not exclude any participants for performance reasons. A detailed description of each site and the way children were recruited can be found below.

@@

Akure is situated in the Ondo State in southwest Nigeria. About 800,000 people live in this medium-sized city. Most inhabitants moved from the nearby urban and rural areas to Akure.

There are two different seasons in Akure's tropical climate: the wet season, which runs from April to October, and the dry season, which runs from November to March. Rich vegetation, including deep forests and productive agriculture, is what defines the city.

@@

Over centuries, Akure has cultivated a sense of unity among its residents, transcending ethnic boundaries and nurturing a strong community spirit grounded in mutual respect and cooperation. This communal ethos extends from familial bonds – encompassing both nuclear and extended families – to broader social networks within neighborhoods, workplaces, and religious congregations. Akure's evolution from a traditional Yoruba kingdom to a thriving urban metropolis is deeply embedded in history and has a rich cultural legacy. The landscape is dotted with evidence of old kingdoms and communities, highlighting the city's lasting influence and crucial role in the history of the area.

@@

The variety of architectural styles and types of homes found in the city are reflected in the housing landscape. Akure offers a diverse range of housing options, including modern urban dwellings scattered across the city and ancient mud-brick homes tucked away in rural hamlets. These housing options can accommodate nuclear families, extended kin networks, and communal living arrangements.

A set of fundamental principles that are deeply embedded in the fabric of Akure society define parenting there. Parents are essential in passing on cultural legacy and forming the identities of future generations because they emphasize reverence for elders, community cohesion, and the value of education. Cultural festivals, ceremonies, and rite of passage serve as platforms for imparting values and fostering social cohesion within the community.

@@

The median age in the Auckland region was 34.7 years. 181,194 Māori people with a median age of 24.9 years were counted in the 2018 census. Overall the population is composed of 53.5% European, 28.2% Asian, 15.5% Pacific Peoples, 11.5% Māori, 2.3% Middle Eastern / Latin American / African and 1.1% Other ethnicity. 92.7% are English speakers, with 4.4% speaking Northern Chinese (Mandarin Chinese), 4.4% speaking Samoan and 2.4% speaking te reo Māori.

Māori peoples arrived in Auckland | Tāmaki Makaurau around 1350 from other Pacific Islands. Before the arrival of European New Zealanders | Pākehā New Zealand was populated by about 20,000 Māori. Aotearoa was colonized by the Dutch and British Empires in the 18th Century. The effects the British Monarchy had on Aotearoa have been at the center of long lasting societal disputes. The Treaty of Waitangi | Te Tiriti o Waitangi is a document signed by 240 Māori chiefs and representatives of the British Crown in 1840 trying to establish a cooperative relationship while redeeming some of the mistakes that have been made. This document is still important for the revitalization of Māori culture throughout Aotearoa.

51.9% of the people in Auckland are employed full time, with an unemployment rate of 4.1% and 30.4% not in the labor force (i.e. retirees). The main occupations are professionals (25.9%) and managers (18.1%) with only 5.9% operating machinery and 7.9% laborers. The people of Auckland are part of a service society with a median income of 34,400NZD and 19.8% of the population earning more than 70,000NZD.

Auckland is made up of mainly sub-urban areas with single family housing. 45.4% of the population own or partly own a house. People mostly live with their core family in loosely separated areas of the city. Those areas commonly have an economical center with shops and restaurants as well as a mix of private and public education institutions.  The socialization goals and cultural norms are comparable to most western, democratic societies. High education standards are deemed as vital for success. Success is determined by career paths, income and general success in professional and private aspects of life. Sport plays a central role in most aspects of society. The role of sports is viewed as essential for living a healthy and successful life and as such is a central part in education programs.

Education levels are generally high in Auckland. 18.6% of the population hold a Bachelor's degree. Interestingly, about 9.1% move overseas for their secondary education. Parenting and childcare practices are comparable to other western cultures, with some adjustments being implemented in recent years to include more Māori cultural practices in education programs. Such as starting the day or specific activities | hui with Karakia, a form of incantation to ensure a favorable outcome to important events and undertakings. (see Literature for further information on early childhood education in New Zealand.)

@@

For this study we reached out to early childcare centers and schools. Two early childcare centers and one primary school agree to participate. Some additional children were tested during home visit sessions. All three educational institutions are visited by children from different cultural and socio-economic backgrounds. One early childcare center is affiliated with the University of Auckland and prioritizes children from University students and staff in their admission process. Testing sessions took place in separated spaces (office or break rooms) with one participant at a time. Participants received a certificate “for becoming a junior scientist” and a small gift (coloring book or sticker set) after their participation. (All numbers in the above description are from the 2018 census data report.)

@@

Both BaYaka and Bandongo populations live (seasonally) in a remote village located in the Likouala region of the Republic of the Congo. The Republic of the Congo is located in the central-western part of sub-Saharan Africa along the equator. The Likouala department is located in the country's north, in the Congo River Basin, bordering the Central African Republic and the Democratic Republic of Congo.
The region's climate is tropical, with temperatures ranging from 34°C to 23°C yearly and 148.73 mm of rainfall yearly (Weather and Climate, 2024). The dry season starts from December until March, with a mild, rainy season from April until August. Heavier rains happen between September and November.

The village is inhabited by approximately 300 BaYaka and 300 Bandongo people (estimates from 2019). The BaYaka and Bandongo people have co-inhabited the rainforests of the Congo basin for approximately 3.000 years (Takeuchi, 2014) and have developed inter-ethnic kin (or patronage) and commercial relationships with one another (Lewis, 2002). Despite this historical coexistence, each cultural group has distinct means of subsistence and speaks their own language; for the Bandongo, this is Bondongo (Bantu C142); for the BaYaka, this is Yaka (C104); although Lingala (Bantu C36d) is often used as lingua-franca between groups) (Hammarström, 2019). The BaYaka are one of many hunter-gatherer societies living in the Congo Basin (Bahuchet, 2017). Like many other hunter-gatherer groups in the world, the BaYaka are egalitarian (Boyette & Hewlett, 2017; Woodburn, 1982), follow an autonomy foundational schema (Boyette & Hewlett, 2017), share resources and parental care (Lewis, 2016) and are characterized by an immediate-return economy with an absence of food storage techniques (Lewis, 2002). The BaYaka forage for food items in the forests they inhabit while supplementing their diets with some agricultural items (Boyette et al., 2020; Kitanishi, 1995). The Bandongo mainly rely on fishing and agriculture to obtain their nutrition (Komatsu, 1998). Although the Bandongo also practice hunting to some extent, it is mainly directed to smaller game species, often sourced with the intent to sell, while BaYaka focus on a wider range of species and bigger game for subsistence (Fa et al., 2016). BaYaka and Bandongo groups sometimes collaborate in their subsistence activities, such as hunting and fishing (Jang & Boyette, 2021; Kandza et al., 2023). Contrary to BaYaka, Bandongo live in a stratified and hierarchical society characterized by wealth accumulation and authority (Boyette et al., 2018).

In the forest camps, the BaYaka inhabit temporary houses (called mongulo or ekuta, depending on the region) built with Marantaceae leaves, lianas, palm tree leaves, etc. (see Lewis, 2002). These structures are occupied by several family members (usually, but not always, grandparents, parents, and their children). Older children may sometimes build their own ekuta, where they sleep alone or with other children. The area of forest camps varies, but usually they are small, leading to ekutas being built relatively close to each other (Lew-Levy et al., 2020). In the village, BaYaka tend to live in unoccupied Bandongo houses or build their own houses with branches and adobe. Bandongo families build permanent habitation with very clear delimitations, usually consisting of a house where the family lives and a kitchen, both built with adobe and thatch roofs (Boyette et al., 2018). Households in the village are usually organized based on patrilineal families.

@@

Beijing, a municipality directly under the Central Government, is the capital of the People's Republic of China and the country's political, economic, cultural and transportation center as well as a famous historic city and a modern international city. Beijing covers an area of 16 410 54 km^2^ with a population of 21 843 000 in 2022. Beijing ranked second in Chinese cities in terms of GDP in 2022. Geographically, Beijing is located in the northern part of the North China Plain.

@@

95% of the sampled Children were recruited from local public kindergarten and primary school. We firstly introduced the experiment to the heads of the kindergarten/school. Once getting their permission, we asked the teachers to help send the printed letter (information about the study, informed consent, as well as the five questions) to the parents. And a Mandarin speaking research assistant then tested kids whose parents signed the informed consent in a separate room in the kindergarten/school. A few of the youngest kids (5% of total) were recruited by directly introducing the experiment to the parents who we personally knew and tested the kids either in a separate room at our institute or at their home.

@@

Moreover, the city is the commercial, financial, and industrial hub of Argentina and a major site for tourism. Buenos Aires has a market economy and the main economic activities center on the service sector (e.g. finance, commerce, tourism, social services) and, to some extent, manufacturing. The city contributes about 20% to the Gross Domestic Product of Argentina. Most people in Buenos Aires live in apartment buildings, but informal settlements have been growing over the last decades. Argentina has experienced one of the highest inflation rates in the world in the past years and poverty has been steadily increasing—though poverty rates in the city of Buenos Aires tend to be lower than in the surrounding Buenos Aires Province.

Commonly, children grow up in nuclear families, but ties with extended family tend to be strong. Generally, family and closeness to family are important cultural values in Argentina. Traditionally, female caregivers had a central role in the family, but this has been undergoing change in recent years, particularly, in middle-class families, with male caregivers taking on caring responsibilities, too. Argentina was the first country in South America to legalize same-sex marriage in 2010 and Buenos Aires in particular is perceived as progressive and liberal city. Some parenting values in middle-class families in Argentina, such as being responsive to the child and providing stimulating encounters, are similar to countries such as the U.S. or France. Caregivers tend to control their children’s schedules/activities, but children do not necessarily perceive this control as negative and caregivers are more permissive with younger children.

@@

### Chimfunshi (Sambia) - Location

Chimfunshi wildlife orphanage is an NGO providing a home for about 150 chimpanzees. The sanctuary is located in the Copperbelt region in the northern part of Zambia close to the DRC border. The chimpanzees are housed in eight different enclosures; each of an area up to 121.5ha. The Sanctuary provides education, housing and medical care for the employees and their families which are in total about 300 people living in 5 different compounds around Chimfunshi.

@@

The country, with approximately 18 million inhabitants, consists of a diverse population primarily composed of Bantu-speaking groups. Among these, the Bemba ethnicity stands out as the largest, accompanied by other significant groups such as Chewa, Tonga, and Lozi. While English serves as the official language, it is spoken natively by only around 2% of the population. However, for those with higher education levels, English becomes the predominant second language. The patterns of migrations has contributed to a rich tapestry of linguistic and cultural diversity across the nation., Due to a robust mining industry in the Copperbelt region, the population hailing from various corners of Zambia and even neighboring countries.

@@

The Hai||om (i.e., treesleepers), also referred to as Heikum or Hai//om, speak Hai||om, which belongs to the Khoekhoe language family often recognized for containing distinct consonants (i.e., “click” sounds). Local dialects (i.e., =Ākhoe) differ across communities. They traditionally lived in small, semi-nomadic bands, loosely organized around kinship ties and shared leadership roles. Like many other forager communities of southern Africa, the Hai||om value egalitarianism, autonomy, and sharing on demand as foundational schemas organizing their social structure.

@@

Due to their historical marginalization and remote location, access to formal education remains challenging for many Hai||om communities. In one community, children have access to a primary school from age 6 onward, whereas in the other community, children need to travel about 35km to access the next school. In both communities, dropout rates are reported low by teachers, likely due to multiple reasons, such as cultural mismatches between often authoritarian teaching styles and foreign learning content at schools, children’s autonomy in deciding on their learning environments, and secondary schools being located far away from their home communities. Traditional knowledge transfer remains vital, with elders playing a significant role in childcare and education. At the same time, children spend considerable amounts of their daily lives in mixed-age peer groups, which present important learning contexts for these children. Children learn through practical experience and participation in daily activities alongside adults and peers.

@@

Exposure to technology, particularly smartphones and other touchscreen devices, is generally limited among the Khwe of Bwabwata also due to their remote location, economic hardships, and lack of access to electricity. However, similar to the Hai||om, the level of access can vary depending on proximity to towns and settlements. Some individuals, especially younger generations, may have some exposure through interactions with external communities or shared devices.

@@

### Malatya (Turkey) - Location

Malatya is a big city located in the Eastern Anatolia region of Turkey with a population of approximately 800 000. Due to its strong economy based on agriculture, it is one of the biggest cities in eastern Turkey.

@@

The urban area of Malatya consists of apartment buildings. Household composition usually involves families with children and multi-generational households. Children live both in nuclear families (with parents and siblings) and in extended families (nuclear family and a set of grandparents). Parental socialization in Malatya aligns with broader Turkish cultural norms, and typically emphasize the importance of children’s interdependence, value of family. Education and learning are seen as important aspects of children’s lives. Parental education levels are generally above the average of Turkey; with most adults have high school or university degrees.

@@

Nyabyeya Forestry College is situated in the Masindi district of western Uganda, approximately 25 km from Masindi Town and 240 km from Kampala, the capital city. The college is adjacent to Budongo Tropical High Forest and 8 km from Budongo conservation field station, where primatologists study wild-ranging chimpanzees. The college draws its sample from the surrounding villages, including Marram, Kyempunu, Nyabigoma, Nyabyeya 1 and 2 and Trading Center, Nyakafunjo, and Kyamasuka.

Nyabyeya is located on a green, gently sloping area behind Nyabyeya Hill and is characterized by tropical vegetation and a savanna climate, characterized by daytime temperatures reaching 32°C/90°F and cooler nights at 17°C. The region observes two distinct agricultural seasons - a wet season from late March to June and August to November, followed by a dry season from early December to March.

The studied villages are culturally very diverse and house ethnic groups such as Bantu, Luo, Nilotics, and Nilo-Hamites. Spoken languages depend on the tribe and include Alur, Lugbara, Runyoro, lendu, kakwa, Acholi, Langi, Luganda, Logo, Kebu, madi, and many others. Kiswahili and English often serve as a common ground. For this study, children who understand Kiswahili were invited.

@@

The community predominantly practices subsistence farming, cultivating crops like maize, cassava, sweet potatoes, and beans. Wealthier families sometimes keep goats, pigs, chickens, or ducks. Additionally, sugarcane and coffee cultivation contribute significantly to the local economy, with the Kinyara Sugar Company playing a pivotal role in sugarcane harvesting.

Living conditions in the community are characterized by extended families residing together, with an average of 10 people under one roof. While temporary round mud huts are common, permanent houses are found in the trading center of Nyabyeya. Most people in the community live by hand to mouth, and parents (especially women) work hard to provide the basics for their household members.

Parental socialization goals vary, with limited guidance for children in some instances, leading to very independent lives at a young age. Cultural practices include the belief in producing many children as a sign of respect, with specific tribes attaching wealth to the number of daughters. Twins are highly honored, and rituals are performed to welcome them. Often, young children take care of their siblings and carry them around on their backs. Children start working early, for example, by helping their parents with farming or cow herding. Often, families cannot afford school materials for their children and struggle to provide regular food. Early marriages and teenage pregnancies often result in girls leaving school earlier.

Parental education levels are generally low, with many parents above 50 having only primary education. The education system faces challenges such as inadequate school supplies, leading to high dropout rates. Children start primary school at ages 5-7, and secondary education runs for four years, with options to pursue advanced levels.

The community exhibits low exposure to technology, with smartphones and touchscreen devices considered luxuries. Limited internet usage and power supply challenges contribute to this low adoption. However, technology is more prevalent in institutions like Nyabyeya Forestry College.
Participating families were recruited by the local research assistants, for example, via personal contact, visiting them at home, or calling them. Families were invited to the Nyabyeya Forestry College (transport offered via boda boda (motorcycle)), and children participated in a quiet room in the College.

@@

Ocuilan is a municipality located in the centre south of Mexico with a population of 36,223 inhabitants, distributed in 48 towns in an area of 344.84 km^2^  inside the Mexican State (Ocuilan, n.d.). Geographically Ocuilan has an altitude of between 1000 to 2 600 m situated 80 km south of Mexico City. (INEGI,2010)

>>>>

---
title: "A universal of human social cognition: Children from 17 communities process gaze in similar ways"
subtitle: "Supplemental material"
author:
  - Manuel Bohn
  - Julia Prein
  - Agnes Ayikoru
  - Florian M. Bednarski
  - Ardain Dzabatou
  - Michael C. Frank 
  - Annette M. E. Henderson 
  - Joan Isabella
  - Josefine Kalbitz 
  - Patricia Kanngiesser
  - Dilara Keşşafoğlu
  - Bahar Köymen
  - Maira V. Manrique-Hernandez 
  - Shirley Magazi
  - Lizbeth Mújica-Manrique
  - Julia Ohlendorf
  - Damilola Olaoba
  - Wesley R. Pieters
  - Sarah Pope-Caldwell
  - Katie Slocombe
  - Robert Z. Sparks
  - Jahnavi Sunderarajan
  - Wilson Vieira
  - Zhen Zhang
  - Yufei Zong
  - Roman Stengelin
  - Daniel B. M. Haun
output: 
  bookdown::pdf_document2:
    toc: yes 
    number_sections: false    
  bookdown::word_document2:
    toc: yes 
    number_sections: false
  bookdown::html_document2:
    toc: yes
    toc_float: true
    fig_caption: yes
    #code_folding: hide
    number_sections: false
bibliography: ../paper/library.bib
csl: science.csl
header-includes:
  \usepackage{caption}
  \renewcommand{\figurename}{Supplementary Figure}
  \renewcommand{\tablename}{Supplementary Table}
---

@@

@@

A total of `r sum(as.numeric(demtab$N))` children between `r round(min(data$ageinyears), 2)` and `r round(max(data$ageinyears), 2)` provided data for the study. Children lived in `r n_distinct(data$community)` different communities, located in `r n_distinct(demtab$country)` different countries. Table \@ref(tab:demtable) gives the sample size per community together with some basic demographic information. The recruitment strategy for each community is reported in the respective site descriptions. Figure \@ref(fig:figpred1)A shows the age distribution in each setting. For some children, the exact birthday was unknown. In such cases, we set the birthday to the 30th of June of the year that would make them fall into the reported age category.

Data from children was only included in the study when they contributed at least four valid test trials. We also excluded the data from children with a diagnosed developmental disorder. In sum, in addition to the sample size reported above, 74 additional children participated in the study but did not contribute data. The main reasons for exclusion were: contribution of less than four valid test trials, technical failures, and missing or implausible demographic information (e.g., when the number of children living in the household was reported to be larger than the household itself or when the number of children reported to live in the household equaled the number of children younger than the child being tested). We did not exclude any participants for performance reasons. A detailed description of each site and the way children were recruited can be found in the respective site description.

@@

Akure is situated in the Ondo State in southwest Nigeria. About 800 000 people live in this medium-sized city. Most inhabitants moved from the nearby urban and rural areas to Akure.

There are two different seasons in Akure's tropical climate: the wet season, which runs from April to October, and the dry season, which runs from November to March. Rich vegetation, including deep forests and productive agriculture, defines the city.

@@

Over centuries, Akure has cultivated a sense of unity among its residents, transcending ethnic boundaries and nurturing a strong community spirit grounded in mutual respect and cooperation. This communal ethos extends from familial bonds – encompassing both nuclear and extended families – to broader social networks within neighborhoods, workplaces, and religious congregations. Akure's evolution from a traditional Yoruba kingdom to a thriving urban metropolis is deeply embedded in history and has a rich cultural legacy, highlighting the city's lasting influence and crucial role in the history of the area.

@@

Akure offers a diverse range of housing options, including modern urban dwellings scattered across the city and ancient mud-brick homes tucked away in rural hamlets. These housing options can accommodate nuclear families, extended kin networks, and communal living arrangements.

Parents are essential in passing on cultural legacy and forming the identities of future generations because they emphasize reverence for elders, community cohesion, and the value of education. Cultural festivals, ceremonies, and rite of passage serve as platforms for imparting values and fostering social cohesion within the community.

@@

The median age in the Auckland region was 34.7 years. 181 194 Māori people with a median age of 24.9 years were counted in the 2018 census. Overall the population is composed of 53.5% European, 28.2% Asian, 15.5% Pacific Peoples, 11.5% Māori, 2.3% Middle Eastern / Latin American / African and 1.1% Other ethnicity. 92.7% are English speakers, with 4.4% speaking Northern Chinese (Mandarin Chinese), 4.4% speaking Samoan and 2.4% speaking te reo Māori.

Māori peoples arrived in Auckland | Tāmaki Makaurau around 1350 from other Pacific Islands. Before the arrival of European New Zealanders | Pākehā New Zealand was populated by about 20 000 Māori. Aotearoa was colonized by the Dutch and British Empires in the 18th Century. The effects the British Monarchy had on Aotearoa have been at the center of long lasting societal disputes. The Treaty of Waitangi | Te Tiriti o Waitangi is a document signed by 240 Māori chiefs and representatives of the British Crown in 1840 trying to establish a cooperative relationship while redeeming some of the mistakes that have been made. This document is still important for the revitalization of Māori culture throughout Aotearoa.

51.9% of the people in Auckland are employed full time, with an unemployment rate of 4.1% and 30.4% retirees. The main occupations are professionals (25.9%) and managers (18.1%) with only 5.9% operating machinery and 7.9% laborers. The people of Auckland are part of a service society with a median income of 34 400NZD and 19.8% of the population earning more than 70 000NZD.

Auckland is made up of mainly sub-urban areas with single family housing. 45.4% of the population own or partly own a house. People mostly live with their core family in loosely separated areas of the city. Those areas commonly have an economical center with shops and restaurants as well as a mix of private and public education institutions. The socialization goals and cultural norms are comparable to most western, democratic societies. High education standards are deemed as vital for success. Success is determined by career paths, income and general success in professional and private aspects of life. The role of sports is viewed as essential for living a healthy and successful life and as such is a central part in education programs.

Education levels are generally high in Auckland. 18.6% of the population hold a Bachelor's degree. Interestingly, about 9.1% move overseas for their secondary education. Parenting and childcare practices are comparable to other western cultures, with some adjustments being implemented in recent years to include more Māori cultural practices in education programs. Such as starting the day or specific activities | hui with Karakia, a form of incantation to ensure a favorable outcome to important events and undertakings (see Literature for further information on early childhood education in New Zealand).

@@

For this study we reached out to early childcare centers and schools. Two early childcare centers and one primary school agreed to participate. Some additional children were tested during home visit sessions. All three educational institutions are visited by children from different cultural and socio-economic backgrounds. One early childcare center is affiliated with the University of Auckland and prioritizes children from University students and staff in their admission process. Testing sessions took place in separated spaces (office or break rooms) with one participant at a time. Participants received a certificate “for becoming a junior scientist” and a small gift (coloring book or sticker set) after their participation. (All numbers in the above description are from the 2018 census data report.)

@@

Both BaYaka and Bandongo populations live (seasonally) in a remote village located in the Likouala region of the Republic of the Congo. The Republic of the Congo is located in the central-western part of sub-Saharan Africa along the equator. The Likouala department is located in the country's north, in the Congo River Basin, bordering the Central African Republic and the Democratic Republic of Congo.
The region's climate is tropical, with temperatures ranging from 23°C to 34°C yearly and 148.73 mm of rainfall yearly (Weather and Climate, 2024). The dry season starts from December until March, with a mild, rainy season from April until August. Heavier rains happen between September and November.

The village is inhabited by approximately 300 BaYaka and 300 Bandongo people (estimates from 2019). The BaYaka and Bandongo people have co-inhabited the rain forests of the Congo basin for approximately 3.000 years (Takeuchi, 2014) and have developed inter-ethnic kin (or patronage) and commercial relationships with one another (Lewis, 2002). Despite this historical coexistence, each cultural group has distinct means of subsistence and speaks their own language; for the Bandongo, this is Bondongo (Bantu C142); for the BaYaka, this is Yaka (C104); although Lingala (Bantu C36d) is often used as lingua-franca between groups) (Hammarström, 2019). The BaYaka are one of many hunter-gatherer societies living in the Congo Basin (Bahuchet, 2017). Like many other hunter-gatherer groups in the world, the BaYaka are egalitarian (Boyette & Hewlett, 2017; Woodburn, 1982), follow an autonomy foundational schema (Boyette & Hewlett, 2017), share resources and parental care (Lewis, 2016) and are characterized by an immediate-return economy with an absence of food storage techniques (Lewis, 2002). The BaYaka forage for food items in the forests they inhabit while supplementing their diets with some agricultural items (Boyette et al., 2020; Kitanishi, 1995). The Bandongo mainly rely on fishing and agriculture to obtain their nutrition (Komatsu, 1998). Although the Bandongo also practice hunting to some extent, it is mainly directed to smaller game species, often sourced with the intent to sell, while BaYaka focus on a wider range of species and bigger game for subsistence (Fa et al., 2016). BaYaka and Bandongo groups sometimes collaborate in their subsistence activities, such as hunting and fishing (Jang & Boyette, 2021; Kandza et al., 2023). Contrary to BaYaka, Bandongo live in a stratified and hierarchical society characterized by wealth accumulation and authority (Boyette et al., 2018).

In the forest camps, the BaYaka inhabit temporary houses (called mongulo or ekuta, depending on the region) built with Marantaceae leaves, lianas, palm tree leaves, etc. (see Lewis, 2002). These structures are occupied by several family members (usually, but not always, grandparents, parents, and their children). Older children may sometimes build their own ekuta, where they sleep alone or with other children. The area of forest camps varies, but usually they are small, leading to ekutas being built relatively close to each other (Lew-Levy et al., 2020). In the village, BaYaka tend to live in unoccupied Bandongo houses or build their own houses with branches and adobe. Bandongo families build permanent habitation with very clear delimitation, usually consisting of a house where the family lives and a kitchen, both built with adobe and thatch roofs (Boyette et al., 2018). Households in the village are usually organized based on patrilineal families.

@@

Beijing, a municipality directly under the Central Government, is the capital of the People's Republic of China and the country's political, economic, cultural and transportation center. Beijing covers an area of 16 410 54 km^2^ with a population of 21 843 000 in 2022. Beijing ranked second in Chinese cities in terms of GDP in 2022. Geographically, Beijing is located in the northern part of the North China Plain.

@@

95% of the sampled children were recruited from local public kindergarten and primary school. We firstly introduced the experiment to the heads of the kindergarten/school. Once getting their permission, we asked the teachers to help send the printed letter (information about the study, informed consent, as well as the five questions) to the parents. A Mandarin speaking research assistant then tested kids whose parents signed the informed consent in a separate room in the kindergarten/school. A few of the youngest kids (5% of total) were recruited by directly introducing the experiment to the parents who we personally knew and tested the kids either in a separate room at our institute or at their home.

@@

Moreover, the city is the commercial, financial, and industrial hub of Argentina and a major site for tourism. Buenos Aires has a market economy and the main economic activities center on the service sector (e.g. finance, commerce, tourism, social services) and, to some extent, manufacturing. The city contributes about 20% to the Gross Domestic Product of Argentina. Most people in Buenos Aires live in apartment buildings, but informal settlements have been growing over the last decades. Argentina has experienced one of the highest inflation rates in the world in the past years and poverty has been steadily increasing — though poverty rates in the city of Buenos Aires tend to be lower than in the surrounding Buenos Aires Province.

Commonly, children grow up in nuclear families, but ties with extended family tend to be strong. Generally, family and closeness to family are important cultural values in Argentina. Traditionally, female caregivers had a central role in the family, but this has been undergoing change in recent years, particularly, in middle-class families, with male caregivers taking on caring responsibilities, too. Argentina was the first country in South America to legalize same-sex marriage in 2010 and Buenos Aires in particular is perceived as progressive and liberal city. Some parenting values in middle-class families in Argentina, such as being responsive to the child and providing stimulating encounters, are similar to western countries. Caregivers tend to control their children’s schedules/activities, but children do not necessarily perceive this control as negative and caregivers are more permissive with younger children.

@@

### Chimfunshi (Zambia) - Location

Chimfunshi wildlife orphanage is an NGO providing a home for about 150 chimpanzees. The sanctuary is located in the Copperbelt region in the northern part of Zambia close to the DRC border. The Sanctuary provides education, housing and medical care for the employees and their families which are in total about 300 people living in 5 different compounds around Chimfunshi.

@@

The country, with approximately 18 million inhabitants, consists of a diverse population primarily composed of Bantu-speaking groups. Among these, the Bemba ethnicity stands out as the largest, accompanied by other significant groups such as Chewa, Tonga, and Lozi. While English serves as the official language, it is spoken natively by only around 2% of the population. However, for those with higher education levels, English becomes the predominant second language. The patterns of migrations have contributed to a rich tapestry of linguistic and cultural diversity across the nation. Due to a robust mining industry in the Copperbelt region, the population hails from various corners of Zambia and even neighboring countries.

@@

The Hai||om (i.e., treesleepers), also referred to as Heikum or Hai//om, speak Hai||om, which belongs to the Khoekhoe language family often recognized for containing distinct consonants (i.e., “click” sounds). Local dialects (i.e., =Ākhoe) differ across communities. They traditionally lived in small, semi-nomadic bands, loosely organized around kinship ties and shared leadership roles.

@@

Due to their historical marginalization and remote location, access to formal education remains challenging for many Hai||om communities. In one community, children have access to a primary school from age 6 onward, whereas in the other community, children need to travel about 35km to access the next school. In both communities, school attendance rates are reported low by teachers, likely due to multiple reasons, such as cultural mismatches between often authoritarian teaching styles and foreign learning content at schools, children’s autonomy in deciding on their learning environments, and secondary schools being located far away from their home communities. Traditional knowledge transfer remains vital, with elders playing a significant role in childcare and education. At the same time, children spend considerable amounts of their daily lives in mixed-age peer groups, which present important learning contexts for these children. Children learn through practical experience and participation in daily activities alongside adults and peers.

@@

Exposure to technology, particularly smartphones and other touchscreen devices, is generally limited among the Khwe of Bwabwata, also due to their remote location, economic hardships, and lack of access to electricity. However, similar to the Hai||om, the level of access can vary depending on proximity to towns and settlements. Some individuals, especially younger generations, may have some exposure through interactions with external communities or shared devices.

@@

### Malatya (Türkiye) - Location

Malatya is a big city located in the Eastern Anatolia region of Türkiye with a population of approximately 800 000. Due to its strong economy based on agriculture, it is one of the biggest cities in eastern Türkiye.

@@

The urban area of Malatya consists of apartment buildings. Household composition usually involves families with children and multi-generational households. Children live both in nuclear families (with parents and siblings) and in extended families (nuclear family and a set of grandparents). Parental socialization in Malatya aligns with broader Turkish cultural norms, and typically emphasize the importance of children’s interdependence and the value of family. Education and learning are seen as important aspects of children’s lives. Parental education levels are generally above the average of Türkiye; with most adults have high school or university degrees.

@@

Nyabyeya Forestry College is situated in the Masindi district of western Uganda, approximately 25 km from Masindi Town and 240 km from Kampala, the capital city. The college is adjacent to Budongo Forest and 8 km from Budongo Conservation Field Station, where primatologists study habituated wild chimpanzees. The sample were invited to participate at the college from the surrounding villages, including Marram, Kyempunu, Nyabigoma, Nyabyeya 1 and 2 and Trading Center, Nyakafunjo, and Kyamasuka.

Nyabyeya village is characterized by tropical vegetation and a savanna climate, characterized by daytime temperatures reaching 32°C and cooler nights at 17°C. The region observes two distinct agricultural seasons - a wet season from late March to June and August to November, followed by a dry season from early December to March.

The studied villages are culturally very diverse and are home to different ethnic groups including Bantu, Luo, Nilotics, and Nilo-Hamites. Spoken languages depend on the tribe and include Alur, Lugbara, Runyoro, Lendu, Kakwa, Acholi, Langi, Luganda, Logo, Kebu, Madi, and many others. Kiswahili and English often serve as a common ground. For this study, children who understood Kiswahili were invited.

@@

The local communities predominantly practices subsistence farming, cultivating crops like maize, cassava, sweet potatoes, and beans. Wealthier families sometimes keep goats, pigs, chickens, or ducks. Additionally, sugarcane and coffee cultivation contribute significantly to the local economy, with the Kinyara Sugar Company playing a pivotal role in sugarcane harvesting. Salaried work is rare and many households struggle to provide essentials, including food, health care, and school materials for family members.

Living conditions in the community are characterized by extended families residing together, with an average of 10 people in each household. Households tend to inhabit several buildings within a compound, with buildings made from traditional materials (wood and mud walls; straw roof) or bricks with metal roofing.

Parental socialization goals vary, but a previous sample of mothers from these communities valued relational socialization goals over autonomous goals (Holden et al., 2022). Cultural practices include the belief in producing many children as a sign of respect, with specific tribes attaching wealth to the number of daughters. Twins are highly honored, and rituals are performed to welcome them. Often, young children take care of their siblings and carry them around on their backs. Children start to help their parents with chores from an early age (collecting fire wood or water, tending crops or animal care).

School is not mandatory and parental education levels in these communities is  generally low, with a minority of mothers in the Holden et al (2022) sample having any secondary level education. The education system faces challenges such as overcrowding and lack of resources, leading to high dropout rates. Children start primary school at ages 5-9, and secondary education runs for four years, with options to pursue advanced levels. Lack of sanitary products to allow attendance of school during menstruation, early marriages and teenage pregnancies often result in girls performing worse than boys and leaving school earlier.

The community exhibits low exposure to technology, with smartphones and touchscreen devices considered luxuries. Limited internet access and power supply challenges contribute to this low adoption. However, technology is more prevalent in institutions like Nyabyeya Forestry College. Participating families were recruited by the local research assistants, for example, via personal contact, visiting them at home, or calling them. Families were invited to participate at the Nyabyeya Forestry College (transport offered via boda boda (motorcycle taxi)), and children participated in a quiet room in the College.

@@

Ocuilan is a municipality located in the center south of Mexico with a population of 36 223 inhabitants, distributed in 48 towns in an area of 344.84 km^2^ inside the Mexican State (Ocuilan, n.d.). Geographically, Ocuilan has an altitude of between 1000 to 2600 m situated 80 km south of Mexico City (INEGI, 2010).

The communities in Ocuilan are distributed on a mountain range called Trans-Mexican volcanic Belt surrounded by valleys and mountains, the climate is semi-humid with summer rains, and average temperatures vary between 8°C and 22°C throughout the year. The higher elevations have mixed pine forests with more tropical vegetation in the lowlands. The distribution of the land is 31.91% agriculture, 0.48% urban zone and the rest forest (INEGI,2010).
